---
title: "Text Analysis: 'The Fellowship of the Ring'"
author: "Larissa Neilson"
date: "2/21/2021"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE)

library(tidyverse)
library(tidytext)
library(textdata)
library(pdftools)
library(ggwordcloud)
library(pacman)
library(grid)
library(plyr)
library(dplyr)
```

```{r, cache = TRUE}
# Import pdf text & wrangle
fotr_text <- pdf_text("fotr.pdf")

fotr_tidy <- data.frame(fotr_text) %>% 
  mutate(text_full = str_split(fotr_text, pattern = '\\n')) %>% 
  unnest(text_full) %>% 
  mutate(text_full = str_trim(text_full)) 

fotr_df <- fotr_tidy %>% 
  slice(-(1:831)) %>% 
  mutate(chapter = case_when(
    str_detect(text_full, "Chapter") ~ text_full,
    TRUE ~ NA_character_
  )) %>% 
  fill(chapter) %>% 
  separate(col = chapter, into = c("ch", "no"), sep = " ") %>% 
  mutate(chapter = as.numeric(no))
```

```{r}
# Get tokens
fotr_tokens <- fotr_df %>% 
  unnest_tokens(word, text_full) %>% 
  select(-fotr_text)

fotr_wordcount <- fotr_tokens %>% 
  dplyr::count(chapter, word)
```

```{r}
fotr_nonstop_words <- fotr_tokens %>% 
  anti_join(stop_words)

nonstop_counts <- fotr_nonstop_words %>% 
  dplyr::count(chapter, word)
```

### Most frequent words

```{r}
# Wrangle to get top 5 words
top_5_words <- nonstop_counts %>% 
  group_by(chapter) %>% 
  arrange(-n) %>% 
  slice(1:5) %>% 
  drop_na()

# Graph it
ggplot(data = top_5_words, aes(x = fct_reorder(word, n), y = n)) +
  geom_col(fill = "lightblue", color = "lightblue4") +
  facet_wrap(~chapter, scales = "free") +
  coord_flip() +
  labs(x = NULL, y = "Word count") +
  ggtitle("Top 5 words") +
  theme_minimal()
```

**Figure 1.** Top five words in each chapter of *The Fellowship of the Ring*. "Frodo" is almost always the most frequent word of each chapter.

```{r, include = FALSE}
# Word clouds! (not to be included)
ch1_top100 <- nonstop_counts %>% 
  filter(chapter == 1) %>% 
  arrange(-n) %>% 
  slice(1:100)

ch1_cloud <- ggplot(data = ch1_top100, aes(label = word)) +
  geom_text_wordcloud(aes(color = n, size = n), shape = "diamond") +
  scale_size_area(max_size = 6) +
  scale_color_gradientn(colors = c("darkgreen","blue","purple")) +
  theme_minimal()

ch1_cloud

ch12_top100 <- nonstop_counts %>% 
  filter(chapter == 12) %>% 
  arrange(-n) %>% 
  slice(1:100)

ch12_cloud <- ggplot(data = ch12_top100, aes(label = word)) +
  geom_text_wordcloud(aes(color = n, size = n), shape = "diamond") +
  scale_size_area(max_size = 6) +
  scale_color_gradientn(colors = c("darkgreen","blue","purple")) +
  theme_minimal()

ch12_cloud
```

```{r}
# Fix to remove duplicate words
top100 <- nonstop_counts %>%
  group_by(chapter) %>%
  plyr::ddply("word", numcolwise(sum)) %>%  
  arrange(-n) %>% 
  slice(1:100)

top100_words <- top100 %>% 
  select(word, n)

# Read in image
ring <- png::readPNG("ring.png")

# Top 100 words!
top100_cloud <- ggplot(data = top100_words, aes(label = word)) +
  theme(panel.background = element_rect(fill = "black")) +
  annotation_custom(rasterGrob(ring, 
                               width = unit(1,"npc"), 
                               height = unit(1,"npc")), 
                               -Inf, Inf, -Inf, Inf) +
  geom_text_wordcloud(aes(color = n, size = n), shape = "circle") +
  scale_size_area(max_size = 6) +
  scale_color_gradientn(colors = c("orange","red","darkred"))
  

top100_cloud
```

**Figure 2.** Top 100 words in the *Fellowship of the Ring* text. "Frodo" is the most common word in the book (n = 1028) which aligns with the results from Figure 1.

### Sentiment Analysis

```{r}
fotr_afinn <- fotr_nonstop_words %>% 
  inner_join(get_sentiments("afinn"))

afinn_counts <- fotr_afinn %>% 
  dplyr::count(chapter, value)

# Find the mean afinn score by chapter: 
afinn_means <- fotr_afinn %>% 
  dplyr::group_by(chapter) %>% 
  dplyr::summarize(mean_afinn = mean(value))

ggplot(data = afinn_means, 
       aes(x = fct_rev(as.factor(chapter)), 
           y = mean_afinn)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(x = "Chapter", y = "Mean afinn value") +
  ggtitle("Sentiment analysis using the afinn lexicon") +
  theme_minimal()
```

**Figure 3.** Sentiment analysis with the AFINN lexicon. Mean AFINN values for each of the 12 chapters. The majority of the words in this book are associated with a negative mean AFINN value, indicating that most words are negative to some extent.

```{r, include = FALSE}
fotr_nrc <- fotr_nonstop_words %>% 
  inner_join(get_sentiments("nrc"))

fotr_nrc_counts <- fotr_nrc %>% 
  dplyr::count(chapter, sentiment)

ggplot(data = fotr_nrc_counts, aes(x = sentiment, y = n)) +
  geom_col() +
  facet_wrap(~chapter) +
  coord_flip()
```

#### Text citation:

Tolkien, J. R. R. (1991). *The Fellowship of the Ring*. HarperCollins.
